Thermodynamic Structure of the Solar Corona: Tomographic Reconstructions and MHD Modeling

### 2020-03-11 response_to_review_SOLA-D-20-00015_v1.txt

-----------------------------------------------------------------------------------------
Q8. The introduction does not have many references to works beyond DEMT and AWSoM. The 2nd paragraph of the introduction offers a chance to remedy this, where you discuss the quiet-Sun corona. I know there are very few works that address the quiet corona in the context of DEM analysis, but you can reference https://www.aanda.org/articles/aa/full_html/2014/04/aa23054-13/aa23054-13.html. You can also mention https://iopscience.iop.org/article/10.1088/0004-637X/795/2/111 here (although you do reference this paper in more detail in the discussion). Another relevant paper which can be referenced, with DEM analysis of a large amount of quiet coronal data over long time periods, is https://advances.sciencemag.org/content/3/7/e1602056 .

A8. We agree it is relevant to expand the references in the Introduction to other DEM studies of the quiet-Sun. In the Introduction, we included now the other references you provided, as well as another one (new text in red). We kept our reference to Hahn & Savin (2014) in the discussion.

-----------------------------------------------------------------------------------------
Q9. Some sections can benefit from weeding out of repetitions. You tend to repeat statements regarding the method a lot. For example:

A9: We agree, and answer below point-by-point.

- Paragraph starting page 13, line 41. This has already been said in the method.

Agreed. We deleted that paragraph moved the detail on the leg-averaged temperature <Tm> right before Table 1 (text in red).

- Sentence Page 14, line 25 'For each leg…

Agreed, removed.

- statement in parenthesis at page 15, line 24.

Agreed, removed.

- Section 4, first two paragraphs.

Agreed, removed. We edited the new first paragraph in this section (red text) following also your suggestions in Q10 below, although we wanted to keep references to specific Figures and Tables here.

- Page 26, line 26:
 'In the streamer belt region we combined the DEMT results and the AWSom magnetic model to infer the energy input phi at the coronal base required to maintain stable coronal loops. We found characteristic values in the range…'
Shorten to: 'The energy input phi at the coronal base, required to maintain stable coronal loops, is in the range…'

Agreed, done.

- Page 26, lines 38-43, between 'Results of the AWSoM….summarized as follows.' can be omitted.

Agreed, done. A little edit was introduced (red text) to start the paragraph.

@@

-----------------------------------------------------------------------------------------
Q10. Minor grammatical improvements, typos, and removal of redundant statements

A10. Thanks very much for this detailed corrections. We agreed with basically all of them and addressed them all. Changes highlighted in red.

### 2020-03-11 lloveras_2020_v2.tex
Albert está trabajando en esto, no tocar todavía, gracias.

{The computational domain of the solar corona} extends from 1 to {$24\,\mrsun$}. \del{The spherical grid in the solar corona is radially stretched (with a maximum resolution {of $0.001\,\mrsun$} near the Sun)} \edit{The spherical grid has an adaptive grid that has fine resolution near the Sun, and increases outward} with the z-axis aligned with the rotation axis in Heliographic Rotation coordinates. The Adaptive Mesh refinement (AMR) resolves the angular cell size to $1.4\mdeg$ between $1 - 1.7\,\mrsun$ and to $2.8\mdeg$ outside this radius range. The solar corona component uses about 3 million cells on $6 \times 8 \times 8$ grid blocks (for CR-2208) and $6 \times 4 \times 4$ grid blocks for CR-2082. Local time stepping is used to speed up the steady state convergence. \edit{The AWSoM simulation of the solar corona for solar minimum conditions represented by CR-2082 and CR-2208 are compared to the DEMT in the results section.}

@@

\noindent
where $h \equiv r - 1\,\mrsun$ is the coronal height measured from the photosphere. In the electron density fit, $\l\,[\mrsun]$ is the density scale height and $N_0\,[\cminvc]$ is the electron density at the footpoint ($h=0$) of the loop. In the electron temperature fit, $a\,[\MK/\mrsun]$ is the slope and $T_0\,[\MK]$ is the electron temperature at the footpoint of the loop. The slope $a$ estimates the radial gradient of the electron temperature along the loop, which we denote as $a = \dr\Tm$ hereafter, \edit{with} $\dr\equiv\er\cdot\nabla$ \edit{being} the radial derivative operator \edit{and} $\er$ is the heliocentric radial unit vector.

@@

{To determine if the leg of a traced field line is} of type up or down, we first determine the Pearson correlation coefficient $\rhoTr$ between the DEMT temperature $\Tm$ and the heliocentric height $r$. We then select field lines for which the temperature {is significantly correlated with height} by requiring $|\rhoTr| > 0.5$. {To test the goodness of the fit \del{to the temperature-height data} we perform a chi-squared test \citep{recipes} considering the uncertainty of the DEMT data, selecting legs for which the fit matches the data with a 90\% confidence level.} In this way, legs for which the DEMT temperature does not show a significant {correlation} with height, or the linear fit to temperature is \edit{weak}, are {excluded from the} analysis. {The test is also applied to the density-height data points, to ensure the trend is reasonably represented by the exponential fit.} Finally, selected legs are then classified as up or down according to $\dr\Tm > 0$ or $\dr\Tm < 0$, respectively. The linear fit allows characterization of the variation of $\Tm$ with height by means of a characteristic temperature gradient $\dr \Tm\,[\MK/\mrsun]$ along each leg. {The chi-squared test to evaluate the quality of the \edit{fit} considers the uncertainty level in the DEMT products due to systematic sources (radiometric calibration and tomographic regularization), that \citet{lloveras_2017} estimated to be $\Delta\Tm\approx 10\%$ and $\Delta\sqravgN\approx 5\%$.} In summary, \edit{a selected leg must meet three conditions:}

@@

\del{To characterize the DEMT results in different magnetic structures, we traced $\sqravgN$ and $\Tm$ along the magnetic field lines of the AWSoM model. For both rotations, all legs that meet the criteria listed in Section \ref{trace} were selected. For each leg, the data points of electron density and electron mean temperature as a function of height were fitted to the Equations \ref{Nfit} and \ref{Tfit}. {As a result, the electron density $N_0(r=1.0\,\mrsun)$ and scale height $\lN$ were computed for each leg, as well as the temperature gradient $\dr\Tm$, and the height-averaged (along the leg) {electron temperature} $\aTm$.}}

@@

For both rotations, the top panels of Figure \ref{rpoint_demt} show the latitude-longitude location (at heliocentric height $r=1.105\,\mrsun$) of all traced field line legs for which criterion (i) of Section \ref{trace} is met. Open legs are indicated in gray color and closed ones in black color. \del{For each leg, the fits to tomografic temperature and density were applied, as given by Equations \ref{Nfit}-\ref{Tfit}.} Considering the DEMT data points and the resulting fits along each leg, the \edit{middle} panels of Figure \ref{rpoint_demt} show the latitude-longitude location of the subset for which also both criteria (ii) and (iii) of Section \ref{trace} are met. Using a four-color code, type 0, I, II and III legs are shown in blue, \edit{orange, red, and cyan} color, respectively. \edit{The bottom panels show histograms of the latitude distribution of the legs of the middle panel, using the same color code.} Of the {$\approx 44000$} legs selected for CR-2082, 20\% are type 0, 31\% are type I, 23\% are type II and 26\% type III. On the other hand, of the \edit{$\approx 50000$} legs selected for CR-2208, 10\% are type 0, \edit{38}\% are type I, \edit{27}\% are type II and \edit{25}\% type III.

@@

{For both rotations, Figure \ref{histos_fulldemt} shows, in a statistical fashion, the DEMT results traced along field lines discriminated by leg type}. From top to bottom {results are shown for} type 0 to type III {legs}{, respectively}. From left to right the panels show the statistical distribution of the electron density value $\NCB \equiv \sqravgN(r=1.055)$ \del{(at the lowest height where the AWSoM results are consistent with coronal conditions)}, $\lN$ {and} $\aTm$, {with the median value $m$ indicated in each plot}.

@@

{Several details from Figure \ref{perf_lat} are \edit{worth highlighted}. Firstly, at most latitudes the overall agreement of the electron density of both models is within $\approx 20\%$ for CR-2082, and $\approx 5\%$ for CR-2208. The noticeable exception is to be found near the open/closed boundaries of both target rotations, where the disagreement between both models can be up to twice as much. In the case of the electron temperature, for both rotations the models agree within $\approx 15\%$ at all latitudes \del{of the streamer belt, showing larger discrepancies in the CHs}. Secondly, for both rotations, and for both models, these plots clearly show the relatively lower temperatures characterizing the low-latitudes of the equatorial streamer belt compared to its mid-latitudes. Thirdly, for both rotations, the latitude of the open/closed magnetic boundary in both hemispheres matches the location of the strongest latitudinal gradient of the DEMT electron density. Note this is not the case for the AWSoM model, that shows a minimum density at the open/closed boundary. \edit{Lastly, the DEMT} electron density decreases from the open/closed boundary towards the poles (in both hemispheres of the two rotations), while the AWSoM model shows the opposite trend. {For comparison, \edit{$B_r$ in the CHs increases} from the open/closed boundary towards the poles for CR-2082, while showing local maxima around latitudes $-75\mdeg$ and $+70\mdeg$ in the case of CR-2208.}}

@@

\section{{Discussion and Conclusions}}\label{discu} 
\del{
We carried out the 3D DEMT reconstruction and MHD modeling of the inner solar corona for two target rotations, CR-2082 and CR-2208. The former belongs to the deep minimum epoch between SCs 23 and 24, while the latter belongs to the end of the declining phase of SC 24. The present work introduces two improvements in the implementation of DEMT, namely the use of only off-limb data in the EUV images, and the inclusion of a 3D regularization scheme (Section \ref{demt}). The MHD model used is here the latest version of the AWSoM component of the SWMF suite  (Section \ref{awsom}).
Based on the magnetic field of AWSoM model and the results of the DEMT model for electron density and temperature, we classified coronal structures in four different types. {Small loop legs} in the core of the streamer belt are classified in down/up types according to their temperature decreasing/increasing with height (dubbed type 0/I). {Large closed legs} of type up that form the envelope of the streamer belt (dubbed type II), and open field lines of type up populating the CHs (dubbed type III).}

\del{Concerning the DEMT reconstructions, Figure \ref{rpoint_demt} shows that, for both rotations,} \edit{Magnetic} field lines of type 0, I and II \edit{were selected to be associated} with increasingly outer {layers} of the equatorial streamer belt. \del{Figure \ref{histos_fulldemt} and Table \ref{tabla_demt} show in 3D quantitative detail how} \edit{These progressively} outer magnetic structures exhibit decreasing coronal base density, increasing density scale height, and increasing electron temperature. For both rotations we find that down legs populate the low latitudes of the streamer belt, while up legs dominate its mid-latitudes. Also, in the case of CR-2082 the fraction of down legs is significantly larger than for CR-2208. These findings are consistent with previous studies by \citet{huang_2012} and \citet{nuevo_2013}. In the case of the latter, they include in their analysis target CR-2081, which is a rotation almost identical to our target CR-2082. Our results for target CR-2082 compare very well with those of \citet{nuevo_2015} and \citet{lloveras_2017} for target CR-2081. As our study uses the improved version of the DEMT technique, such comparison provided a consistency \edit{check.} \del{for the changes implemented in the DEMT technique.} \edit{For both rotations, type III field lines in the CHs are characterized by sub-MK temperatures, and electron density values of order $\approx 1/2$ of those observed for the type 0 and type I lines in the core of the equatorial streamer.}

{\del{In the streamer belt region we combined the DEMT results and the AWSoM magnetic model to infer the energy input flux $\phi_h$ at the coronal base required to maintain stable coronal loops. We found characteristic values in the range \\}
\edit{The energy input flux $\phi_h$ at the coronal base, required to maintain stable coronal loops, is in the range} $\phi_h\approx 0.5-1.5 \times 10^5\,\erg\,\cminvs\,\s^{-1}$, depending on the rotation and the type of {loop, matching the values reported by} \citet{maccormack_2017}. Based on spectroscopic data of the EIS instrument in quiet-Sun regions \citet{hahn_2014} showed that, if the observed non-thermal broadening are assigned to Alfvén waves, their energy flux at the coronal base is estimated to be in the range $\approx 1.5-2.5\times 10^5\,\erg\,\cminvs\,\s^{-1}$. A large fraction of the coronal base energy input flux $\phi_h$ estimated in this work, or even its totality, could then be accounted for by Alfvén waves (see the discussion in \citealt{maccormack_2017}).}

\del{
Results of the AWSoM model {were compared} to those of the DEMT reconstructions {in the range of coronal heights simultaneously covered by both models, namely $\approx 1.05-1.20\,\mrsun$.} In its current implementation, the AWSoM model can not reproduce down loops (type 0), so the comparison is restricted to type I, II and III field lines.} The detailed comparison shown in this work can be summarized as follows. For CR-2082, the electron density of both models agree within $\approx 20\%$ {in most regions}, while for CR-2208 the agreement is within $\approx 5\%$. {The noticeable exception is to be found near the open/closed boundaries of both target rotations, where the disagreement between both models can be up to twice as much.} In the case of the electron temperature, both models agree within $\approx 10-15\%$ in all regions, for both rotations\del{, with larger discrepancies in the CHs.} This level of agreement between both models {(within or slightly beyond the uncertainty level of DEMT results)} is considerably better than that reported in previous works. \citet{jin_2012} and \citet{oran_2015}{, who used previous versions of the AWSoM model,} reported electron density values of the AWSoM model {differing by $\approx 50\%$ compared to the DEMT reconstructions, both in the equatorial streamer and CH regions.}

@@

\del{Within the quantitative match between both models summarized in the previous paragraph, there are some qualitative agreements and {disagreements worth noticing}. }For both rotations, the AWSoM model reproduces the relatively lower temperatures found by DEMT to characterize the low-latitudes of the equatorial streamer belt compared to its mid-latitudes. On the other hand, while the latitude of the open/closed magnetic boundary in both hemispheres matches the location of the strongest latitudinal gradient of the DEMT electron density {(physically expected in transitioning from magnetically closed to open regions)}, this is not the case for the AWSoM model, that shows a minimum density at the open/closed boundary. Also, while the DEMT electron density decreases from the open/closed boundary towards the poles (in both hemispheres of the two rotations), {as expected in transitioning from the source region of the slow to the fast component of the solar wind}, the AWSoM model shows the opposite trend. This behavior is notoriously opposite to that reported in the AWSoM model version used by \citet{oran_2015}, in which the electron density decreases from the open/closed boundary towards the poles. These unphysical characteristics of the results of the AWSoM model in the \edit{range} of low heights analysed here ($r\lesssim 1.2\,\mrsun$), may be attributed to the less reliable values of $B_r$ provided by both the GONG and ADAPT-GONG maps at subpolar latitudes. This will \edit{be investigated} in a follow up article focusing on the current deep minimum epoch, during which the large-scale corona shows the simplest possible structure.

>>>>

{The computational domain of the solar corona} extends from 1 to {$24\,\mrsun$}. \edit{The spherical grid has an adaptive grid that has fine resolution near the Sun, and increases outward} with the z-axis aligned with the rotation axis in Heliographic Rotation coordinates. The Adaptive Mesh refinement (AMR) resolves the angular cell size to $1.4\mdeg$ between $1 - 1.7\,\mrsun$ and to $2.8\mdeg$ outside this radius range. The solar corona component uses about 3 million cells on $6 \times 8 \times 8$ grid blocks (for CR-2208) and $6 \times 4 \times 4$ grid blocks for CR-2082. Local time stepping is used to speed up the steady state convergence. \edit{The AWSoM simulation of the solar corona for solar minimum conditions represented by CR-2082 and CR-2208 are compared to the DEMT in the results section.}

@@

\noindent
where $h \equiv r - 1\,\mrsun$ is the coronal height measured from the photosphere. In the electron density fit, $\l\,[\mrsun]$ is the density scale height and $N_0\,[\cminvc]$ is the electron density at the footpoint ($h=0$) of the loop. In the electron temperature fit, $a\,[\MK/\mrsun]$ is the slope and $T_0\,[\MK]$ is the electron temperature at the footpoint of the loop. The slope $a$ estimates the radial gradient of the electron temperature along the loop, which we denote as $a = \dr\Tm$ hereafter, \edit{with} $\dr\equiv\er\cdot\nabla$ \edit{being} the radial derivative operator \edit{and} $\er$ the heliocentric radial unit vector.

@@

{To determine if the leg of a traced field line is} of type up or down, we first determine the Pearson correlation coefficient $\rhoTr$ between the DEMT temperature $\Tm$ and the heliocentric height $r$. We then select field lines for which the temperature {is significantly correlated with height} by requiring $|\rhoTr| > 0.5$. {To test the goodness of the fit we perform a chi-squared test \citep{recipes} considering the uncertainty of the DEMT data, selecting legs for which the fit matches the data with a 90\% confidence level.} In this way, legs for which the DEMT temperature does not show a significant {correlation} with height, or the linear fit to temperature is \edit{weak}, are {excluded from the} analysis. {The test is also applied to the density-height data points, to ensure the trend is reasonably represented by the exponential fit.} Finally, selected legs are then classified as up or down according to $\dr\Tm > 0$ or $\dr\Tm < 0$, respectively. The linear fit allows characterization of the variation of $\Tm$ with height by means of a characteristic temperature gradient $\dr \Tm\,[\MK/\mrsun]$ along each leg. {The chi-squared test to evaluate the quality of the \edit{fit} considers the uncertainty level in the DEMT products due to systematic sources (radiometric calibration and tomographic regularization), that \citet{lloveras_2017} estimated to be $\Delta\Tm\approx 10\%$ and $\Delta\sqravgN\approx 5\%$.} In summary, \edit{a selected leg must meet three conditions:}

@@

@@

For both rotations, the top panels of Figure \ref{rpoint_demt} show the latitude-longitude location (at heliocentric height $r=1.105\,\mrsun$) of all traced field line legs for which criterion (i) of Section \ref{trace} is met. Open legs are indicated in gray color and closed ones in black color. Considering the DEMT data points and the resulting fits along each leg, the \edit{middle} panels of Figure \ref{rpoint_demt} show the latitude-longitude location of the subset for which also both criteria (ii) and (iii) of Section \ref{trace} are met. Using a four-color code, type 0, I, II and III legs are shown in blue, \edit{orange, red, and cyan} color, respectively. \edit{The bottom panels show histograms of the latitude distribution of the legs of the middle panel, using the same color code.} Of the {$\approx 44000$} legs selected for CR-2082, 20\% are type 0, 31\% are type I, 23\% are type II and 26\% type III. On the other hand, of the \edit{$\approx 50000$} legs selected for CR-2208, 10\% are type 0, \edit{38}\% are type I, \edit{27}\% are type II and \edit{25}\% type III.

@@

{For both rotations, Figure \ref{histos_fulldemt} shows, in a statistical fashion, the DEMT results traced along field lines discriminated by leg type}. From top to bottom {results are shown for} type 0 to type III {legs}{, respectively}. From left to right the panels show the statistical distribution of the electron density value $\NCB \equiv \sqravgN(r=1.055)$, \edit{the scale height $\lN$, and the height-averaged (along the leg) electron temperature $\aTm$}, {with the median value $m$ indicated in each plot}.

@@

{Several details from Figure \ref{perf_lat} are \edit{worth highlighted}. Firstly, at most latitudes the overall agreement of the electron density of both models is within $\approx 20\%$ for CR-2082, and $\approx 5\%$ for CR-2208. The noticeable exception is to be found near the open/closed boundaries of both target rotations, where the disagreement between both models can be up to twice as much. In the case of the electron temperature, for both rotations the models agree within $\approx 15\%$ at all latitudes. Secondly, for both rotations, and for both models, these plots clearly show the relatively lower temperatures characterizing the low-latitudes of the equatorial streamer belt compared to its mid-latitudes. Thirdly, for both rotations, the latitude of the open/closed magnetic boundary in both hemispheres matches the location of the strongest latitudinal gradient of the DEMT electron density. Note this is not the case for the AWSoM model, that shows a minimum density at the open/closed boundary. \edit{Lastly, the DEMT} electron density decreases from the open/closed boundary towards the poles (in both hemispheres of the two rotations), while the AWSoM model shows the opposite trend. {For comparison, \edit{$B_r$ in the CHs increases} from the open/closed boundary towards the poles for CR-2082, while showing local maxima around latitudes $-75\mdeg$ and $+70\mdeg$ in the case of CR-2208.}}

@@

\section{{Discussion and Conclusions}}\label{discu}

\edit{Magnetic field lines of type 0, I and II were selected to be associated with increasingly outer layers of the equatorial streamer belt (Figure \ref{rpoint_demt}). These magnetic structures progressively exhibit decreasing coronal base density, increasing density scale height, and increasing electron temperature, as informed in 3D quantitative detail in Figure \ref{histos_fulldemt} and Table \ref{tabla_demt}.} For both rotations we find that down legs populate the low latitudes of the streamer belt, while up legs dominate its mid-latitudes. Also, in the case of CR-2082 the fraction of down legs is significantly larger than for CR-2208. These findings are consistent with previous studies by \citet{huang_2012} and \citet{nuevo_2013}. In the case of the latter, they include in their analysis target CR-2081, which is a rotation almost identical to our target CR-2082. Our results for target CR-2082 compare very well with those of \citet{nuevo_2015} and \citet{lloveras_2017} for target CR-2081. As our study uses the improved version of the DEMT technique, such comparison provided a consistency \edit{check.} \edit{For both rotations, type III field lines in the CHs are characterized by sub-MK temperatures, and electron density values of order $\approx 1/2$ of those observed for the type 0 and type I lines in the core of the equatorial streamer.}

\edit{The energy input flux $\phi_h$ at the coronal base, required to maintain stable coronal loops, is in the range} $\phi_h\approx 0.5-1.5 \times 10^5\,\erg\,\cminvs\,\s^{-1}$, depending on the rotation and the type of {loop, matching the values reported by} \citet{maccormack_2017}. Based on spectroscopic data of the EIS instrument in quiet-Sun regions \citet{hahn_2014} showed that, if the observed non-thermal broadening are assigned to Alfvén waves, their energy flux at the coronal base is estimated to be in the range $\approx 1.5-2.5\times 10^5\,\erg\,\cminvs\,\s^{-1}$. A large fraction of the coronal base energy input flux $\phi_h$ estimated in this work, or even its totality, could then be accounted for by Alfvén waves (see the discussion in \citealt{maccormack_2017}).

\edit{The comparison of the results of the AWSoM model to the DEMT reconstructions can be summarized as follows.} For CR-2082, the electron density of both models agree within $\approx 20\%$ {in most regions}, while for CR-2208 the agreement is within $\approx 5\%$. {The noticeable exception is to be found near the open/closed boundaries of both target rotations, where the disagreement between both models can be up to twice as much.} In the case of the electron temperature, both models agree within $\approx 10-15\%$. This level of agreement between both models {(within or slightly beyond the uncertainty level of DEMT results)} is considerably better than that reported in previous works. \citet{jin_2012} and \citet{oran_2015}{, who used previous versions of the AWSoM model,} reported electron density values of the AWSoM model {differing by $\approx 50\%$ compared to the DEMT reconstructions, both in the equatorial streamer and CH regions.}

@@

For both rotations, the AWSoM model reproduces the relatively lower temperatures found by DEMT to characterize the low-latitudes of the equatorial streamer belt compared to its mid-latitudes. On the other hand, while the latitude of the open/closed magnetic boundary in both hemispheres matches the location of the strongest latitudinal gradient of the DEMT electron density {(physically expected in transitioning from magnetically closed to open regions)}, this is not the case for the AWSoM model, that shows a minimum density at the open/closed boundary. Also, while the DEMT electron density decreases from the open/closed boundary towards the poles (in both hemispheres of the two rotations), {as expected in transitioning from the source region of the slow to the fast component of the solar wind}, the AWSoM model shows the opposite trend. This behavior is notoriously opposite to that reported in the AWSoM model version used by \citet{oran_2015}, in which the electron density decreases from the open/closed boundary towards the poles. These unphysical characteristics of the results of the AWSoM model in the \edit{range} of low heights analysed here ($r\lesssim 1.2\,\mrsun$), may be attributed to the less reliable values of $B_r$ provided by both the GONG and ADAPT-GONG maps at subpolar latitudes. This will \edit{be investigated} in a follow up article focusing on the current deep minimum epoch, during which the large-scale corona shows the simplest possible structure.
